### 2018-02-16 experiment_info.Rmd
First commit with seed set data

---
title: "Untitled"
output: html_document
---
# Intro

# Methods

The study was done with x species where all the species where consider focal species from xxx to xxx at University of New England. The rooms where temperature controlled with temperature between xxx at day time to xxx at nigth time.

Soil was mixed with sand

Hybrids where avoided

Pot size

Number of individuals (started with 24 ending with 16, lack of space)

# Results

Plots of pollen

Morphometry

Ovules

Seed set

Pollen grwoth

Pic of pollen with SEM (pollen size)

Pollen tubes and number of pollen grains

#Discussion