### 2018-02-20 experiment_info.Rmd
Updating plots

---
title: "Glasshouse experiment"
output: html_document
---
# Intro

# Methods

The study was done with x species (table 1) where all the species where consider focal species from xxx to xxx at University of New England. The rooms where temperature controlled with temperature between xxx at day time to xxx at nigth time.

Flowers were emasculated the day prior anthesis and hand pollinated next day with a toothpick. Had-pollination was realized with 3-4 gentle touches on the surface of the stigma. The mixes of pollen were realized based on the pollen counts on an eppendorf. Pollen on anthers was counted through Neubaeur chamber.

Flower selection criteria: Big flowers, easy/fast growing.

Soil was mixed with sand

Hybrids where avoided

Pot size

Number of individuals (started with 24 ending with 16, lack of space)

# Results

**Ovules**

```{r, echo=FALSE}

#OVYLE NUMBER OF THE DIFFERENT SPECIES N=15

library(reshape2)

# Consider NA's as 0, I have to double check this

ovules <- read.csv("Ovules.csv", sep=";")

#plot(ovules)

mean_ovule <- dcast(Species ~ ., value.var = "Ovules", fun.aggregate = mean, data = ovules, na.rm= TRUE)

colnames(mean_ovule)[2] <- "avg"

sd_ovule <- dcast(Species ~ ., value.var = "Ovules", fun.aggregate = sd, data = ovules, na.rm= TRUE)

colnames(sd_ovule)[2] <- "sdev"

x <- 1:8

plot(x, mean_ovule$avg,
     ylim=range(c(mean_ovule$avg-sd_ovule$sdev, mean_ovule$avg+sd_ovule$sdev)),
     pch=19, xlab="", xaxt='n', ylab="Mean +/- SD",
     main="N of Ovules")
# hack: we draw arrows but with very special "arrowheads"
arrows(x, mean_ovule$avg-sd_ovule$sdev, x, mean_ovule$avg+sd_ovule$sdev, length=0.05, angle=90, code=3)

axis(x, at=1:8, labels = mean_ovule$Species,cex=0.2, tick = FALSE, padj= -0.5, side = 1)

```

**Seed set Petunia**

```{r, echo=FALSE}
library(reshape2)

# Consider NA's as 0, I have to double check this

seed_set <- read.csv("PEIN_seed_set.csv", sep=";")

#Delete Cross with lost seeds
seed_set[c(12,17,19),4] <- NA
#Deleted

mean_seed_seet <- dcast(Treatment ~ ., value.var = "Seed.production", fun.aggregate = mean, data = seed_set, na.rm= TRUE)

mean_seed_seet$.-> avg

sd_seed_seet <- dcast(Treatment ~ ., value.var = "Seed.production", fun.aggregate = sd, data = seed_set, na.rm= TRUE)

sd_seed_seet$. -> sdev

sdev[4] <- 0

#plot(seed_set, cex.axis=0.5)

x <- 1:7

plot(x, avg,
     ylim=range(c(avg-sdev, avg+sdev)),
     pch=19, xlab="", xaxt='n', ylab="Mean +/- SD",
     main="PEIN seed set"
)
# hack: we draw arrows but with very special "arrowheads"
arrows(x, avg-sdev, x, avg+sdev, length=0.05, angle=90, code=3)

axis(x, at=1:7, labels = mean_seed_seet$Treatment,cex.axis=0.8, tick = FALSE, padj= -0.5, side = 1)

```

Plots of pollen

Morphometry

Seed set

Pollen grwoth

Pic of pollen with SEM (pollen size)

Pollen tubes and number of pollen grains

#Discussion