### 2018-03-29 experiment_info.Rmd
a little bit of methods

---
title: "Glasshouse experiment"
output: html_document
---
# Intro

**Paragraph 1**

A common trend in natural ecosystems is pollinator sharing (REF). Therefore, pollen from different species can arrive to the stigmas of other species. The real impact of pollen in a natural ecosystem is still unknown. Nonetheless, some advances has been done with recent works where heterospecific pollen effect has been evaluated (REFS).

**Paragraph 2**

**Paragraph 3**

**Paragraph 4**

We review the classic predicted traits of hterospecific pollen effect through a matrix of intereactions with heterogeneous traits.

# Methods

The study was conducted in a glasshouse at University of New England (Armidale, Australia). Rooms were temperature controlled depending on the requirements of the species. The species selected belonged to three different families, Solanaceae, Brassicaceae and Convolvulaceae. The criteria of species selection was based on close/distant related species (see phylogenetic tree for relatedness fig 1), easy species to grow, low structural flower complexity. For the purpose of the experiment all the species where consider as pollen recipient and as pollen donor (see interaction matrix, fig 2). Species were watered once or twice per day and fertilized weekly (NPK 23: 3.95: 14).

For all the species, the day prior anthesis flowers were emasculated and then hand-pollinated next day with toothpicks with 3-4 gentle touches over the stigma surface.

Check fertiliser type: Hortico (all purpose soluble fertiliser) NPK 23: 3.95: 14

Table of species:

**Solanaceae**

SOLY "Tommy Joe"
CAAN "California Wonder""
PEIN "Petunia integrifolia"
SOME "Solanum melongera"

**Brassicaceae**

Sinapis alba White mustard
Brassica oleracea var. capitata
Brassica rapa var. chinensis

**Convolvulaceae**

Ipomoea purpurea

Possible species to include:

Ipomoea aquatica

I. coccinea

Convolvulus sabatius

Sysimbrium spp

Raphanus spp

Eruca sativa

Flowers were emasculated the day prior anthesis and hand pollinated next day with a toothpick. Had-pollination was realized with 3-4 gentle touches on the surface of the stigma. The mixes of pollen were realized based on the pollen counts on an eppendorf. Pollen on anthers was counted through Neubaeur chamber.

Flower selection criteria: Big flowers, easy/fast growing.

Soil was mixed with sand

Hybrids where avoided

Pot size

Number of individuals (started with 24 ending with 16, lack of space)

# Results

**Ovules**

```{r, echo=FALSE}

#OVYLE NUMBER OF THE DIFFERENT SPECIES N=15

library(reshape2)

# Consider NA's as 0, I have to double check this

ovules <- read.csv("Ovules.csv", sep=";")

#plot(ovules)

mean_ovule <- dcast(Species ~ ., value.var = "Ovules", fun.aggregate = mean, data = ovules, na.rm= TRUE)

colnames(mean_ovule)[2] <- "avg"

sd_ovule <- dcast(Species ~ ., value.var = "Ovules", fun.aggregate = sd, data = ovules, na.rm= TRUE)

colnames(sd_ovule)[2] <- "sdev"

x <- 1:8

plot(x, mean_ovule$avg,
     ylim=range(c(mean_ovule$avg-sd_ovule$sdev, mean_ovule$avg+sd_ovule$sdev)),
     pch=19, xlab="", xaxt='n', ylab="Mean +/- SD",
     main="N of Ovules")
# hack: we draw arrows but with very special "arrowheads"
arrows(x, mean_ovule$avg-sd_ovule$sdev, x, mean_ovule$avg+sd_ovule$sdev, length=0.05, angle=90, code=3)

axis(x, at=1:8, labels = mean_ovule$Species,cex=0.2, tick = FALSE, padj= -0.5, side = 1)

```

**Seed set Petunia**

```{r, echo=FALSE}
library(reshape2)

# Consider NA's as 0, I have to double check this

seed_set <- read.csv("PEIN_seed_set.csv", sep=";")

#Delete Cross with lost seeds
seed_set[c(12,17,19),4] <- NA
#Deleted

mean_seed_seet <- dcast(Treatment ~ ., value.var = "Seed.production", fun.aggregate = mean, data = seed_set, na.rm= TRUE)

mean_seed_seet$.-> avg

sd_seed_seet <- dcast(Treatment ~ ., value.var = "Seed.production", fun.aggregate = sd, data = seed_set, na.rm= TRUE)

sd_seed_seet$. -> sdev

sdev[4] <- 0

#plot(seed_set, cex.axis=0.5)

x <- 1:7

plot(x, avg,
     ylim=range(c(avg-sdev, avg+sdev)),
     pch=19, xlab="", xaxt='n', ylab="Mean +/- SD",
     main="PEIN seed set"
)
# hack: we draw arrows but with very special "arrowheads"
arrows(x, avg-sdev, x, avg+sdev, length=0.05, angle=90, code=3)

axis(x, at=1:7, labels = mean_seed_seet$Treatment,cex.axis=0.8, tick = FALSE, padj= -0.5, side = 1)

```

Plots of pollen

Morphometry

Seed set

Pollen grwoth

Pic of pollen with SEM (pollen size)

Pollen tubes and number of pollen grains

#Discussion