### 2018-10-15 Rmd/Data_visualization_common_names.Rmd
took me too much time BUT I made it

---
title: "Data_visualization"
output:
      pdf_document
---

###SEED SET PER SPECIES WITH DIFFERENT TREATMENTS

```{r,message=FALSE, warning=FALSE, echo=FALSE, cache=FALSE, fig.height=6, fig.width=7,out.width='100%'}
#Preparing seed set plots, starting with Petunia

#There are 3 crosses were seeds were lost (2,7,9)
#load libraries
library(reshape2)
library(ggplot2)
library(dplyr)
library(gridExtra)
library(grid)
library(lattice)
library(ggpubr)
#read data and cleaning
pein_seed_set_final <- read.csv("Data/pein_seed_set_final_common_names.csv")
pein_seed_set_final=pein_seed_set_final[,-1]
#load the palette for all the 4 plots
cbPalette <- c( "#56B4E9","#E69F00", "#999999", "#009E73")
#Colur per family

a <- ggplot(pein_seed_set_final, aes(x = factor(Treatment, levels=unique(Treatment)), y = Seed.production)) +   geom_boxplot(outlier.shape = NA)+
  labs(title="A) Petunia",x="", y = "Seeds")+aes(fill=Family)+theme(axis.text.x =element_blank())+
  theme(plot.title = element_text(hjust = 0.5)) +scale_fill_manual(values=cbPalette)+ geom_jitter(width = 0.3,shape=1,size=0.8, aes(colour=Family))+scale_color_manual(values = cbPalette) + stat_summary(fun.y=mean, geom="point", shape="*", size=5) +theme(legend.position="none")

#
##
###
#C. ANNUUM 
###
##
#
caan_seed_set_final <- read.csv("Data/caan_seed_set_final.csv")
caan_seed_set_final=caan_seed_set_final[,-1]

#Colur per family
#Different colour per family
b <- ggplot(caan_seed_set_final, aes(x = factor(Treatment, levels=unique(Treatment)), y = Seed.production)) +   geom_boxplot(outlier.shape = NA)+
  labs(title="B) Capsicum",x="", y = "")+aes(fill=Family)+theme(axis.text.x =element_blank())+
  theme(plot.title = element_text(hjust = 0.5)) +scale_fill_manual(values=cbPalette)+ geom_jitter(width = 0.3,shape=1,size=0.8, aes(colour=Family))+scale_color_manual(values = cbPalette) + stat_summary(fun.y=mean, geom="point", shape="*", size=5) +theme(legend.position="none")

#
##
###
#S. LYCOPERSICUM
###
##
#

soly_seed_set_final <- read.csv("Data/soly_seed_set_final_common_names.csv")
soly_seed_set_final=soly_seed_set_final[,-1]

cbPalette <- c( "#56B4E9","#E69F00", "#999999", "#009E73")
#Different colour per family
c <- ggplot(soly_seed_set_final, aes(x = factor(Treatment, levels=unique(Treatment)), y = Seed.production)) +   geom_boxplot(outlier.shape = NA)+
  labs(title="C) Tomato",x="", y = "Seeds")+aes(fill=Family)+theme(axis.text.x = element_text(angle = 60, hjust = 1))+
  theme(plot.title = element_text(hjust = 0.5)) +scale_fill_manual(values=cbPalette)+ geom_jitter(width = 0.3,shape=1,size=0.8, aes(colour=Family))+scale_color_manual(values = cbPalette) + stat_summary(fun.y=mean, geom="point", shape="*", size=5) +theme(legend.position="none")

#
##
###
#S. MELONGENA
###
##
#

some_seed_set_final <- read.csv("Data/some_seed_set_final_common_names.csv")
some_seed_set_final=some_seed_set_final[,-1]

#Different colour per family
d <- ggplot(some_seed_set_final, aes(x = factor(Treatment, levels=unique(Treatment)), y = Seed.production)) +   geom_boxplot(outlier.shape = NA)+
  labs(title="D) Eggplant",x="", y = "")+aes(fill=Family)+theme(axis.text.x = element_text(angle = 60, hjust = 1))+
  theme(plot.title = element_text(hjust = 0.5)) +scale_fill_manual(values=cbPalette)+ geom_jitter(width = 0.3,shape=1,size=0.8, aes(colour=Family))+scale_color_manual(values = cbPalette) + stat_summary(fun.y=mean, geom="point", shape="*", size=5) +theme(legend.position="none")

gl <- lapply(list(a,b,c,d), ggplotGrob)
library(grid)
widths <- do.call(unit.pmax, lapply(gl, "[[", "widths"))
heights <- do.call(unit.pmax, lapply(gl, "[[", "heights"))
lg <- lapply(gl, function(g) {g$widths <- widths; g$heights <- heights; g})

grobz <- lapply(list(a, b, c, d), ggplotGrob)
grobz.plot <- arrangeGrob( grobs = list(rbind(grobz[[1]], grobz[[3]], size = "last"),
                                        rbind(grobz[[2]], grobz[[4]], size = "last")), 
                           ncol = 2, top="Solanaceae Focal Species")

grid.draw(grobz.plot)

load("legend.rda")
grid.newpage()
grid.draw(legend) 
```

**Figure 1**. Seed set of Solanaceae species with the different pollination treatments (N=10) A) Petunia, B) Capsicum, C) Tomato, D) Eggplant. In total 9 different crosses with 50% heterospecific pollen appear from left to right. The focal treatment with itself is shown in blank to maintain simmetry among pannels. Moreover, for each species are also shown in grey hand cross pollination, hand self pollination, control (bagged emasculated flowers) and flower control (bagged flowers).

```{r,message=FALSE, warning=FALSE, echo=FALSE, cache=FALSE, fig.height=6, fig.width=7,out.width='100%'}
library(dplyr)

#load libraries
library(reshape2)
library(ggplot2)
library(dplyr)
library(gridExtra)
library(grid)
library(lattice)
library(gtable)
##
###
#B. OLERACEA
###
##
#

brol_seed_set_final <- read.csv("Data/brol_seed_set_final.csv")
brol_seed_set_final=brol_seed_set_final[,-1]
#Different colour per family
a <- ggplot(brol_seed_set_final, aes(x = factor(Treatment, levels=unique(Treatment)), y = Seed.production)) +   geom_boxplot(outlier.shape = NA)+
  labs(title="A) Wild cabagge",x="", y = "Seeds")+aes(fill=Family)+theme(axis.text.x =element_blank())+
  theme(plot.title = element_text(hjust = 0.5)) +scale_fill_manual(values=cbPalette)+ geom_jitter(width = 0.3,shape=1,size=0.8, aes(colour=Family))+scale_color_manual(values = cbPalette) + stat_summary(fun.y=mean, geom="point", shape="*", size=5) +theme(legend.position="none")

#
##
###
#B. RAPA
###
##
#

brra_seed_set_final <- read.csv("Data/brra_seed_set_final.csv")
brra_seed_set_final=brra_seed_set_final[,-1]

#Different colour per family
b <- ggplot(brra_seed_set_final, aes(x = factor(Treatment, levels=unique(Treatment)), y = Seed.production)) +   geom_boxplot(outlier.shape = NA)+
  labs(title="B) Pak choi",x="", y = "Seeds")+aes(fill=Family)+theme(axis.text.x =element_blank())+
  theme(plot.title = element_text(hjust = 0.5)) +scale_fill_manual(values=cbPalette)+ geom_jitter(width = 0.3,shape=1,size=0.8, aes(colour=Family))+scale_color_manual(values = cbPalette) + stat_summary(fun.y=mean, geom="point", shape="*", size=5) +theme(legend.position="none")

#
##
###
#E. VERSICARIA
###
##
#

ersa_seed_set_final <- read.csv("Data/ersa_seed_set_final_common_names.csv")
ersa_seed_set_final=ersa_seed_set_final[,-1]

#Different colour per family
c <- ggplot(ersa_seed_set_final, aes(x = factor(Treatment, levels=unique(Treatment)), y = Seed.production)) +   geom_boxplot(outlier.shape = NA)+
  labs(title="C) Rocket",x="", y = "Seeds")+aes(fill=Family)+theme(axis.text.x = element_text(angle = 60, hjust = 1))+
  theme(plot.title = element_text(hjust = 0.5)) +scale_fill_manual(values=cbPalette)+ geom_jitter(width = 0.3,shape=1,size=0.8, aes(colour=Family))+scale_color_manual(values = cbPalette) + stat_summary(fun.y=mean, geom="point", shape="*", size=5) +theme(legend.position="none")

##
###
#S. ALBA  
###
##
#

sial_seed_set_final <- read.csv("Data/sial_seed_set_final_common_names.csv")
sial_seed_set_final=sial_seed_set_final[,-1]

#Different colour per family
d <- ggplot(sial_seed_set_final, aes(x = factor(Treatment, levels=unique(Treatment)), y = Seed.production)) +   geom_boxplot(outlier.shape = NA)+
  labs(title="D) White mustard",x="", y = "Seeds")+aes(fill=Family)+theme(axis.text.x = element_text(angle = 60, hjust = 1))+
  theme(plot.title = element_text(hjust = 0.5)) +scale_fill_manual(values=cbPalette)+ geom_jitter(width = 0.3,shape=1,size=0.8, aes(colour=Family))+scale_color_manual(values = cbPalette) + stat_summary(fun.y=mean, geom="point", shape="*", size=5) +theme(legend.position="none")

gl <- lapply(list(a,b,c,d), ggplotGrob)
library(grid)
widths <- do.call(unit.pmax, lapply(gl, "[[", "widths"))
heights <- do.call(unit.pmax, lapply(gl, "[[", "heights"))
lg <- lapply(gl, function(g) {g$widths <- widths; g$heights <- heights; g})

grobz <- lapply(list(a, b, c, d), ggplotGrob)
grobz.plot <- arrangeGrob( grobs = list(rbind(grobz[[1]], grobz[[3]], size = "last"),
                                        rbind(grobz[[2]], grobz[[4]], size = "last")), ncol = 2,top="Brassicaceae Focal species")

grid.draw(grobz.plot)

load("legend.rda")
grid.newpage()
grid.draw(legend) 
```

**Figure 2**. Seed set of Brassicaceae species with the different pollination treatments (N=10) A) Wild cabbage, B) Pak choi, C) Rocket, D) White mustard. In total 9 different crosses with 50% heterospecific pollen appear from left to right. The focal treatment with itself is shown in blank to maintain simmetry among pannels. Moreover, for each species are also shown in grey hand cross pollination, hand self pollination, control (bagged emasculated flowers) and flower control (bagged flowers).

\newpage
```{r,message=FALSE, warning=FALSE, echo=FALSE, cache=FALSE, fig.height=3.7, fig.width=7,out.width='100%'}

library(reshape2)
library(ggplot2)
library(dplyr)
library(gridExtra)
library(grid)
library(lattice)

#Preparing seed set plots, starting with Petunia
#
##
###
#I. AQUATICA
###
##
#
ipaq_seed_set_final <- read.csv("Data/ipaq_seed_set_final_common_names.csv")
ipaq_seed_set_final=ipaq_seed_set_final[,-1]

cbPalette <- c( "#56B4E9","#E69F00", "#999999", "#009E73")

#Different colour per family
a <- ggplot(ipaq_seed_set_final, aes(x = factor(Treatment, levels=unique(Treatment)), y = Seed.production)) +   geom_boxplot(outlier.shape = NA)+
  labs(title="A) Water morning glory",x="", y = "Seeds")+aes(fill=Family)+theme(axis.text.x = element_text(angle = 60, hjust = 1))+
  theme(plot.title = element_text(hjust = 0.5)) +scale_fill_manual(values=cbPalette)+ geom_jitter(width = 0.3,shape=1,size=0.8, aes(colour=Family))+scale_color_manual(values = cbPalette) + stat_summary(fun.y=mean, geom="point", shape="*", size=5) +theme(legend.position="none")

#IPPU
ippu_seed_set_final <- read.csv("Data/ippu_seed_set_final_common_names.csv")
ippu_seed_set_final=ippu_seed_set_final[,-1]

#Different colour per family
b <- ggplot(ippu_seed_set_final, aes(x = factor(Treatment, levels=unique(Treatment)), y = Seed.production)) +   geom_boxplot(outlier.shape = NA)+
  labs(title="B) Morning glory",x="", y = "Seeds")+aes(fill=Family)+theme(axis.text.x = element_text(angle = 60, hjust = 1))+
  theme(plot.title = element_text(hjust = 0.5)) +scale_fill_manual(values=cbPalette)+ geom_jitter(width = 0.3,shape=1,size=0.8, aes(colour=Family))+scale_color_manual(values = cbPalette) + stat_summary(fun.y=mean, geom="point", shape="*", size=5) +theme(legend.position="none")

#There are 3 crosses were seeds were lost (2,7,9)
#load libraries

#read data and cleaning

gl <- lapply(list(a,b), ggplotGrob)
library(grid)
widths <- do.call(unit.pmax, lapply(gl, "[[", "widths"))
heights <- do.call(unit.pmax, lapply(gl, "[[", "heights"))
lg <- lapply(gl, function(g) {g$widths <- widths; g$heights <- heights; g})

grobz <- lapply(list(a, b), ggplotGrob)
grobz.plot <- arrangeGrob(grobs = list(rbind(grobz[[1]], size = "last"),
                                        rbind(grobz[[2]], size = "last")), ncol = 2,top="Convolvulaceae")

grid.draw(grobz.plot)

load("legend.rda")
grid.newpage()
grid.draw(legend)

```

**Figure 3**. Seed set of Convolvulaceae species with the different pollination treatments (N=10) A) Water morning glory, B) Morning glory. In total 9 different crosses with 50% heterospecific pollen appear from left to right. The focal treatment with itself is shown in blank to maintain simmetry among pannels. Moreover, for each species are also shown in grey hand cross pollination, hand self pollination, control (bagged emasculated flowers) and flower control (bagged flowers).

\newpage
#CHOLOROPLAST RBCL PHYLOGENY
```{r echo=FALSE, out.width='80%', fig.align='center'}
knitr::include_graphics('Images/all_species.png')
```

**Figure 4**. Evolutionary distances of the ten species used in the experiment using maximum composite likelihood method.

\newpage
###POLLEN MIX RATIOS PER FAMILY -3 NEXT PAGES- (N=3)

Each histogram shows the pollen ratio in percentage of the focal and non-focal species. The barplots are organized per family.

```{r, message=FALSE, warning=FALSE, echo=FALSE,fig.width=8}
pollen_ratios_brassicaceae <- read.csv("Data/pollen_ratios_brassicaceae.csv", header = T)
ggplot(pollen_ratios_brassicaceae, aes(x=spp, y=ratio, fill=variable)) +
  geom_bar(stat='identity', position='dodge') +labs(title="Brassicaceae", x="", y="Pollen mix ratio") + 
  theme(axis.text.x=element_text(angle=60,hjust=1))+geom_hline(yintercept = 50)+
  scale_fill_manual(values=c("grey10", "grey60"))+ ylim(0,100)

```
\newpage
```{r, message=FALSE, warning=FALSE, echo=FALSE,fig.width=8}
pollen_ratios_solanaceae <- read.csv("Data/pollen_ratios_solanaceae.csv", header = T)
ggplot(pollen_ratios_solanaceae, aes(x=spp, y=ratio, fill=variable)) +
  geom_bar(stat='identity', position='dodge') +labs(title="Solanaceae", x="", y="Pollen mix ratio") +
  theme(axis.text.x=element_text(angle=60,hjust=1))+geom_hline(yintercept = 50)+
  scale_fill_manual(values=c("grey10", "grey60"))+ ylim(0,100)

```
\newpage
```{r, message=FALSE, warning=FALSE, echo=FALSE,fig.width=8}
pollen_ratios_convolvulaceae <- read.csv("Data/pollen_ratios_convolvulaceae.csv", header = T)
ggplot(pollen_ratios_convolvulaceae, aes(x=spp, y=ratio, fill=variable)) + 
  geom_bar(stat='identity', position='dodge') +labs(title="Convolvulaceae", x="", y="Pollen mix ratio") + 
  theme(axis.text.x=element_text(angle=60,hjust=1))+geom_hline(yintercept = 50)+
  scale_fill_manual(values=c("grey10", "grey60"))+ ylim(0,100)

```

\newpage
###TOTAL POLLEN ON STIGMA -NEXT 3 PAGES-
```{r, message=FALSE, warning=FALSE, echo=FALSE,fig.width=8}
total_pollen_brassicaceae <- read.csv("Data/total_pollen_brassicaceae.csv", header=T)

ggplot(total_pollen_brassicaceae, aes(x=spp, y=ratio, fill=variable)) +
  geom_bar(stat='identity', position='dodge') +labs(title="Brassicaceae", x="", y="Pollen grains on stigma") + 
  theme(axis.text.x=element_text(angle=60,hjust=1))+
  scale_fill_manual(values=c("grey10", "grey60"))

```

\newpage
```{r, message=FALSE, warning=FALSE, echo=FALSE,fig.width=8}

total_pollen_solanaceae <- read.csv("Data/total_pollen_solanaceae.csv", header=T)

ggplot(total_pollen_solanaceae, aes(x=spp, y=ratio, fill=variable)) +
  geom_bar(stat='identity', position='dodge') +labs(title="Solanaceae", x="", y="Pollen mix ratio") +
  theme(axis.text.x=element_text(angle=60,hjust=1))+
  scale_fill_manual(values=c("grey10", "grey60"))

```
\newpage

```{r, message=FALSE, warning=FALSE, echo=FALSE,fig.width=8}

total_pollen_convolvulaceae <- read.csv("Data/total_pollen_convolvulaceae.csv", header=T)

ggplot(total_pollen_convolvulaceae, aes(x=spp, y=ratio, fill=variable)) + 
  geom_bar(stat='identity', position='dodge') +labs(title="Convolvulaceae", x="", y="Pollen mix ratio") + 
  theme(axis.text.x=element_text(angle=60,hjust=1))+
  scale_fill_manual(values=c("grey10", "grey60"))

```

\newpage
###POLLEN PER ANTHER (N=20, I. aquatica N=10)
```{r,message=FALSE, warning=FALSE, echo=FALSE, cache=FALSE, fig.height=3}
library(ggplot2)

all <- read.csv("Data/all.csv", header=T)

cbPalette <- c( "#56B4E9","#E69F00","#009E73")

#all

 ggplot(all, aes(x = reorder(species, pollen, FUN = median, na.rm = TRUE), y = pollen)) +   geom_boxplot(outlier.shape = NA)+
  labs(title="All species",x="", y = "Pollen per anther")+aes(fill=family)+theme(axis.text.x = element_text(angle = 60, hjust = 1))+
  theme(plot.title = element_text(hjust = 0.5)) +scale_fill_manual(values=cbPalette)+ geom_jitter(width = 0.3,shape=1,size=0.8, aes(colour=family))+scale_color_manual(values = cbPalette) + stat_summary(fun.y=mean, geom="point", shape="*", size=5) +theme(legend.title = element_blank())

#Samte thing but without eggplant to improve visualization

#all

 all <- all[all$species !="Eggplant",]

 ggplot(all, aes(x = reorder(species, pollen, FUN = median, na.rm = TRUE), y = pollen)) +   geom_boxplot(outlier.shape = NA)+
  labs(title="All species(-Eggplant)",x="", y = "Pollen per anther")+aes(fill=family)+theme(axis.text.x = element_text(angle = 60, hjust = 1))+
  theme(plot.title = element_text(hjust = 0.5)) +scale_fill_manual(values=cbPalette)+ geom_jitter(width = 0.3,shape=1,size=0.8, aes(colour=family))+scale_color_manual(values = cbPalette) + stat_summary(fun.y=mean, geom="point", shape="*", size=5) +theme(legend.title = element_blank())

```

\newpage
###OVULES (N=15)
```{r,message=FALSE, warning=FALSE, echo=FALSE, cache=FALSE, fig.height=3}

d <- read.csv("Data/d.csv")

cbPalette <- c( "#56B4E9","#E69F00","#009E73")

ggplot(d, aes(x = reorder(species, ovules, FUN = median, na.rm = TRUE), y = ovules)) +
  geom_boxplot(outlier.shape = NA)+labs(title="All species",x="", y = "Nº of ovules")+
  aes(fill=family)+theme(axis.text.x = element_text(angle = 60, hjust = 1))+
  theme(plot.title = element_text(hjust = 0.5)) +scale_fill_manual(values=cbPalette)+
  geom_jitter(width = 0.3,shape=1,size=0.8, aes(colour=family))+
  scale_color_manual(values = cbPalette) + 
  stat_summary(fun.y=mean, geom="point", shape="*", size=5) +theme(legend.title = element_blank())

 d <- d[d$species !="Eggplant",]

 ggplot(d, aes(x = reorder(species, ovules, FUN = median, na.rm = TRUE), y = ovules)) +
  geom_boxplot(outlier.shape = NA)+labs(title="All species(-Eggplant)",x="", y = "Nº of ovules")+
  aes(fill=family)+theme(axis.text.x = element_text(angle = 60, hjust = 1))+
  theme(plot.title = element_text(hjust = 0.5)) +scale_fill_manual(values=cbPalette)+
  geom_jitter(width = 0.3,shape=1,size=0.8, aes(colour=family))+
  scale_color_manual(values = cbPalette) + 
  stat_summary(fun.y=mean, geom="point", shape="*", size=5) +theme(legend.title = element_blank())
```

\newpage

###POLLEN OVULE RATIO. TOTAL POLLEN PER FLOWER DIVIDED BY NUMBER OF OVULES

```{r,message=FALSE, warning=FALSE, echo=FALSE, cache=FALSE, fig.height=6}
library(dplyr)
library(ggplot2)

tab <- read.csv("Data/tab_common_names.csv", sep="", header = T)

ggplot(tab, aes(x=reorder(species, pollen_ovule_ratio, colour = cut), y=pollen_ovule_ratio))+aes(fill=family)+
  geom_bar(stat="identity", width=0.5)+theme(axis.text.x = element_text(angle = 60, hjust=1))+scale_fill_manual(values=cbPalette)
```

\newpage

###STIGMA AREA. BRASSICACEAE+1 AND SOLANACEAE+1

```{r,message=FALSE, warning=FALSE, echo=FALSE, cache=FALSE, fig.height=6}

t_stigma_area <- read.csv("Data/t_stigma_area.csv")
t_stigma_area_brassicaceae <- read.csv("Data/t_stigma_area_brassicaceae.csv")
t_stigma_area_solanaceae <- read.csv("Data/t_stigma_area_solanaceae.csv")
t_stigma_area_convolvulaceae <- read.csv("Data/t_stigma_area_convolvulaceae.csv")

theme_set(theme_gray())
a <- ggplot(t_stigma_area_brassicaceae, aes(x = um, fill = species_common)) +
  geom_density(alpha = 0.7) +
  theme(legend.position = c(0.8, 0.8))

b <- ggplot(t_stigma_area_solanaceae, aes(x = um, fill = species_common)) +
  geom_density(alpha = 0.7) +
  theme(legend.position = c(0.8, 0.8))

cowplot::plot_grid(a,b, labels = "AUTO")
```

\newpage

###STIGMA AREA. ALL SPECIES, COLOUR PER SPECIES AND FAMILY

I like this type of density plot and I wanted to show the differences of stigma area of our 3 familys. The x axis shows the area in micrometers (um^2).

```{r,message=FALSE, warning=FALSE, echo=FALSE, cache=FALSE, fig.height=6}

t_stigma_area <- read.csv("Data/t_stigma_area.csv")
t_stigma_area_brassicaceae <- read.csv("Data/t_stigma_area_brassicaceae.csv")
t_stigma_area_solanaceae <- read.csv("Data/t_stigma_area_solanaceae.csv")
t_stigma_area_convolvulaceae <- read.csv("Data/t_stigma_area_convolvulaceae.csv")

theme_set(theme_gray())
a <- ggplot(t_stigma_area, aes(x = um, fill = species_common)) +
  geom_density(alpha = 0.7) +
  theme(legend.position = c(0.8, 0.8))

cowplot::plot_grid(a, labels = "AUTO")

###All Area per family

theme_set(theme_gray())
a <- ggplot(t_stigma_area, aes(x = um, fill = family)) +
  geom_density(alpha = 0.7) +
  theme(legend.position = c(0.8, 0.8))

cowplot::plot_grid(a, labels = "AUTO")

```