### 2018-10-23 Manuscript/MS.Rmd
file rename

---
title: "| Título en español \n| Title in English\n"
author: "Name Apellidos^1^, Nombre Apellidos^2^"
csl: ecosistemas.csl
output:
  pdf_document: default
  word_document:
    fig_caption: yes
    fig_height: 6
    fig_width: 7.5
    highlight: null
    reference_docx: Ecosistemas_template.docx
bibliography:
- references.bib
- knitcitations.bib
---

> (1) Dirección del primer autor.

> (2) Dirección del segundo autor.

> Autor para correspondencia: Nombre Apellidos [minombre@micorreo.com]

# Abstract

Pollinator sharing can have negative consequences for species fitness with the arrival of foreign pollen. However, the costs of heterospecific pollen are not yet well understood. For this reason, we have conducted a glasshouse experiment where we try to understand how phylogenetic relatedness and the different traits of these species are involved in this process. We experimentally crossed 10 species belonging to three different families: Brassicaceae, Solanaceae and Convolvulaceae. Overall, more than 4000 crosses were done and seed set and pollen tubes were considered as proxy of effect. We found that for all species foreign pollen (50% or less) reduced seed set. Moreover, the seed set reduction is not dependent on the degree of relatedness of the pollen donor. However, the effect is governed by the degree of relatedness and the traits of the species recipient. Our results show that the outcome of heterospecific pollen deposition is determined in greater degree by the traits of the pollen recipient than the pollen donor and that certain traits such as compatibility system are crucial to understand the costs of heterospecific pollen.

# Keywords

> keyword1; keyword2; keyword3; keyword4

```{r knitcitations, echo=FALSE, cache = FALSE}
library(knitcitations)
cleanbib()   
cite_options(citation_format = "pandoc")
```

# Introduction

Plant ecological generalization means that a wide range of floral visitors will visit this particular species. Normally plant species co-occur with other different species in natural systems and therefore pollinator sharing is common characteristic in natural ecosystems.

# Methods

## Análisis de datos

```{r read_data, eval = FALSE, echo = FALSE}
dataset <- read.csv("mydata.csv")
```

Ajustamos un modelo lineal:

$$
y_{i} = \alpha + \beta*x_{i} 
$$

```{r model, echo = FALSE, eval = FALSE}
model <- lm(y ~ x)
```

Utilizamos R `r citep(citation())` y Rmarkdown `r citep(list(citation("knitr"), citation("rmarkdown")))` para todos nuestros análisis. Para ajustar los modelos mixtos utilizamos `lme4` `r citep(citation("lme4"))`.

# Results

Esta sección está dividida en subsecciones.

## Subsección 1

Los árboles de la parcela A fueron más altos que en la parcela B (altura media: `r mean(25, 31, 28)` vs `r mean(13, 19, 16)` m). Y muchos más resultados que se actualizan dinámicamente.

## Subsección 2

Texto.

## Subsección 3

Texto.

# Discusión

Texto.

# Conclusiones

Texto.

# Agradecimientos

Agradecimientos.

###### References

```{r write_citations, cache=FALSE, include=FALSE}
write.bibtex(file="knitcitations.bib")
```

<div id = "refs"></div>

###### TABLA 1

**Tabla 1**. Cada tabla debe aportar su correspondiente encabezamiento explicativo. En los Artículos de investigación, de revisión y Comunicaciones breves se aportarán los encabezamientos tanto en castellano como en inglés, en letra Arial 10 y en página independiente. Es importante que sean simples y que no superen el ancho una página DIN A4 vertical. Los originales se deben aportar en formato tabla y no en formato imagen.

**Table 1**. Table heading in English.

```{r Tabla1, results='asis', echo=FALSE, cache=FALSE}
library("knitr")
kable(head(iris))
```

###### PIES DE FIGURA

**Figura 1**. Pie de figura 1.

**Figura 2**. Pie de figura 2.

###### FIGURE LEGENDS

**Figure 1**. Figure caption.

**Figure 2**. Figure caption.

###### FIGURA 1

```{r Fig1, echo=FALSE, fig.cap="Figura 1. Esto es un ejemplo.", cache=FALSE}
x <- rnorm(100)
y <- jitter(x, 1000)
plot(x, y)
```

###### FIGURA 2

```{r Fig2, echo=FALSE, fig.cap="Figura 2. Segundo ejemplo", cache=FALSE}
a <- sort(rnorm(100))
b <- c(rep("Group Small", 35), rep("Group Big", 65))
boxplot(a ~ b)
```