### 2019-03-02 Effect_size_plot.Rmd
solanaceae panel, still have to figure out how to fix sizes!

---
title: "Effect_size_Markdown"
output:
  pdf_document: default
  html_document: default
---

```{r, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE,fig.height=4,fig.width=7}

library(effsize)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(cowplot)
library(gridExtra)
library(egg)
library(magrittr)
library(multipanelfigure)

#SOLY
load("Manuscript_draft/effect_size_species/soly_effect_size.RData")
p1<- ggplot(soly_effect_size, aes(Species_1,Cohen_d, size=10)) + theme_bw(base_size=10)+ geom_point(show.legend = FALSE,aes(color=factor(Family))) +geom_errorbar(show.legend=FALSE, aes(x = Species_1, ymin = Lower, ymax = Upper, size=2,color=factor(Family)),
width = 0.2)+scale_color_manual("Family",values=c("#0072B2", "#009E73", "#E69F00", "#D55E00"))+
  scale_fill_manual("Family",values=c("#0072B2", "#009E73", "#E69F00", "#D55E00"))+
  xlab("Treatments") + ylab("Cohen's d") + rotate(ylim = c(-6, 2))+guides(fill=FALSE)+
  geom_hline(yintercept=0, linetype="dashed", color = "black")

#SOME
load("Manuscript_draft/effect_size_species/some_effect_size.RData")
p5<- ggplot(some_effect_size, aes(Species_1,Cohen_d, size=10)) + theme_bw(base_size=10)+geom_point(show.legend = FALSE,aes(color=factor(Family))) +geom_errorbar(show.legend=FALSE, aes(x = Species_1, ymin = Lower, ymax = Upper, size=2,color=factor(Family)),
width = 0.2)+scale_color_manual("Family",values=c("#0072B2", "#009E73", "#E69F00", "#D55E00"))+
  scale_fill_manual("Family",values=c("#0072B2", "#009E73", "#E69F00", "#D55E00"))+
  xlab("Treatments") + ylab("Cohen's d") + rotate(ylim = c(-6, 2))+guides(fill=FALSE)+
  geom_hline(yintercept=0, linetype="dashed", color = "black")

#CAAN
load("Manuscript_draft/effect_size_species/caan_effect_size.RData")
p3<- ggplot(caan_effect_size, aes(Species_1,Cohen_d, size=10)) + theme_bw(base_size=10)+ geom_point(show.legend = FALSE,aes(color=factor(Family))) +geom_errorbar(show.legend=FALSE, aes(x = Species_1, ymin = Lower, ymax = Upper, size=2,color=factor(Family)),
width = 0.2)+scale_color_manual("Family",values=c("#0072B2", "#009E73", "#E69F00", "#D55E00"))+
  scale_fill_manual("Family",values=c("#0072B2", "#009E73", "#E69F00", "#D55E00"))+
  xlab("Treatments") + ylab("Cohen's d") + rotate(ylim = c(-6, 2))+guides(fill=FALSE)+
  geom_hline(yintercept=0, linetype="dashed", color = "black")

#PEIN
load("Manuscript_draft/effect_size_species/pein_effect_size.RData")
p4<- ggplot(pein_effect_size, aes(Species_1,Cohen_d, size=10)) + theme_bw(base_size=10)+ geom_point(show.legend = FALSE,aes(color=factor(Family))) +geom_errorbar(show.legend=FALSE, aes(x = Species_1, ymin = Lower, ymax = Upper, size=2,color=factor(Family)),
width = 0.2)+scale_color_manual("Family",values=c("#0072B2", "#009E73", "#E69F00", "#D55E00"))+
  scale_fill_manual("Family",values=c("#0072B2", "#009E73", "#E69F00", "#D55E00"))+
  xlab("Treatments") + ylab("Cohen's d") + rotate(ylim = c(-6, 2))+guides(fill=FALSE)+
  geom_hline(yintercept=0, linetype="dashed", color = "black")

 grid.arrange(p1, p5, p3, p4, nrow=2,ncol=2,top="Solanaceae")

```