### 2019-03-02 Effect_size_families_plot.Rmd
markdown with first species

---
title: "Effect_size_families"
author: "Jose Barragan"
date: "3 March 2019"
output: html_document
---

```{r, echo=FALSE, message=FALSE, warning=FALSE,cache=FALSE,fig.height=9,fig.width=7}

library(effsize)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(cowplot)
library(gridExtra)
library(egg)
library(magrittr)
library(multipanelfigure)
library(grid)

load("Manuscript_draft/effect_size_species/some_effect_size_family.RData")

p1<- ggplot(soly_effect_size, aes(Family,Cohen_d, size=10)) + theme_bw(base_size=10)
p1 + geom_point(show.legend = FALSE,aes(color=factor(Family))) +
  geom_errorbar(show.legend=FALSE, aes(x = Family, ymin = Lower, 
ymax = Upper, size=2,color=factor(Family)), width = 0.2)+
  scale_color_manual("Family",values=c("#E69F00", "#0072B2", "#009E73", "#D55E00"))+
  scale_fill_manual("Family",values=c("#E69F00", "#0072B2", "#009E73", "#D55E00"))+
  xlab("Treatments") + ylab("Cohen's d") + rotate()+guides(fill=FALSE)+
  geom_hline(yintercept=0, linetype="dashed", color = "black")

```