### 2019-07-10 Manuscript_draft/Protest_plot.Rmd
edits

---
title: "Untitled"
output:
  pdf_document: default
  html_document: default
---

```{r,message=FALSE, warning=FALSE, echo=FALSE, cache=FALSE, fig.height=6}
library(ggplot2)
library(reshape2)
library(ggplot2)
library(dplyr)
library(gridExtra)
library(grid)
library(lattice)
pro_dat<-readRDS("Data/pro_dat.RData")

p1<-ggplot(pro_dat,aes(x=traits,y=pro_cor))+
  geom_bar(stat="identity",position="dodge")+theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
p2<-ggplot(pro_dat,aes(x=traits,y=sum_squ))+
  geom_bar(stat="identity",position="dodge")+theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
p3<-ggplot(pro_dat,aes(x=traits,y=signif))+
  geom_bar(stat="identity",position="dodge")+theme(axis.text.x=element_text(angle=60,hjust=1))

grid.draw(rbind(ggplotGrob(p1),ggplotGrob(p2), ggplotGrob(p3), size="last"))

```

```{r,message=FALSE, warning=FALSE, echo=FALSE, cache=FALSE, fig.height=6}
library(ggplot2)
library(reshape2)
library(ggplot2)
library(dplyr)
library(gridExtra)
library(grid)
library(lattice)
pro_dat_sol<-readRDS("Data/pro_dat_sol.RData")

p1<-ggplot(pro_dat_sol,aes(x=traits_sol,y=pro_cor_sol))+
  geom_bar(stat="identity",position="dodge")+theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
p2<-ggplot(pro_dat_sol,aes(x=traits_sol,y=sum_squ_sol))+
  geom_bar(stat="identity",position="dodge")+theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
p3<-ggplot(pro_dat_sol,aes(x=traits_sol,y=signif_sol))+
  geom_bar(stat="identity",position="dodge")+theme(axis.text.x=element_text(angle=60,hjust=1))

grid.draw(rbind(ggplotGrob(p1),ggplotGrob(p2), ggplotGrob(p3), size="last"))

```

```{r,message=FALSE, warning=FALSE, echo=FALSE, cache=FALSE, fig.height=6}
library(ggplot2)
library(reshape2)
library(ggplot2)
library(dplyr)
library(gridExtra)
library(grid)
library(lattice)
pro_dat_sol<-readRDS("Data/pro_dat_sol.RData")

p1<-ggplot(pro_dat_sol,aes(x=traits_sol,y=pro_cor_sol))+
  geom_bar(stat="identity",position="dodge")+theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
p2<-ggplot(pro_dat_sol,aes(x=traits_sol,y=sum_squ_sol))+
  geom_bar(stat="identity",position="dodge")+theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
p3<-ggplot(pro_dat_sol,aes(x=traits_sol,y=signif_sol))+
  geom_bar(stat="identity",position="dodge")+theme(axis.text.x=element_text(angle=60,hjust=1))

grid.draw(rbind(ggplotGrob(p1),ggplotGrob(p2), ggplotGrob(p3), size="last"))

```