### 2019-09-27 Manuscript_draft/Plot_transitivity_example.Rmd
Trying to adress some comments

---
title: "Untitled"
output:
  pdf_document: default
  html_document: default
---
```{r, echo=FALSE, message=FALSE,warning=FALSE,cache=FALSE,fig.height=5,fig.width=5}
library(igraph)
library(grid)
library(gridExtra)
library(egg)
x <- matrix(1:9, nrow = 3, dimnames = list(c("A","B","C"), c("A","B","C")))
x[1,] <- c(0,1,1) 
x[2,] <- c(0,0,1) 
x[3,] <- c(0,0,0) 
net_2=graph.adjacency(x,mode="directed",weighted=TRUE,diag=FALSE) 
b <- plot.igraph(net_2,vertex.size=35,vertex.label=NA,edge.color="grey4",edge.width=E(net_2)$weight/0.6, edge.arrow.size=1)

```