### 2019-10-17 Manuscript_draft/Figure4.Rmd
Figures 2 to 5 ready for publication. Lacking 1 in High resolution I think I'm going to change the plant images...

---
output:
  pdf_document: default
  html_document: default
---

```{r, echo=FALSE, message=FALSE, warning=FALSE,cache=FALSE,fig.height=2.5,fig.width=6}

library(effsize)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(cowplot)
library(gridExtra)
library(egg)
library(magrittr)
library(multipanelfigure)
library(grid)

load("effect_size_species/effect_size_total.RData")
png("Figure4.png", width = 11000, height = 5000, res = 1500)

 ggplot(all, aes(Family,Cohen_d,colour=Family_1), size=10) + theme_bw(base_size=10)+theme(axis.text.y = element_text(face = "italic"))+ geom_point( aes(colour=Family_1),size=2) +
  geom_errorbar(aes(x = Family, ymin = Lower, 
ymax = Upper,colour=Family_1), width = 0.2, size=0.8, alpha=0.5)+
  scale_color_manual(name  ="Family",values=c("#D55E00", "#009E73", "#0072B2"), labels=c("Brassicaceae","Convolvulaceae","Solanaceae"))+ guides(colour = guide_legend(override.aes = list(size=3)))+guides(colour = guide_legend(override.aes = list(colour=c("#0072B2", "#009E73", "#D55E00"))))+
  xlab("") + ylab("Hedges' g") + rotate()+
  geom_hline(yintercept=0, linetype="dashed", color = "black")+scale_alpha(guide = 'none')
  dev.off()

```