### 2020-04-05 Supplementary_material/Supp_Material.Rmd
Supp material on the way

---
title: "__Supporting Information:__"
output: 
  pdf_document:
    keep_tex: yes
    fig_caption: yes
    toc: false
---

# Recipient and donor characteristics govern hierarchical structure in a heterospecific pollen competition network of co-flowering plants

__Authors:__ Jose B. Lanuza, Ignasi Bartomeus, Tia Lynn Ashman, Romina Rader

The following Supporting Information is available for this article:

__Table S1.__ Species names, common names, varieties and sources of the different seeds.

__Table S2.__ Numerical values of all the traits measured for each species.

__Table S3.__ Seed set in percentage for hand cross-pollination, hand self-pollination, natural selfing and apomixis for all species.

__Table S4.__ Species x species matrix with the significance of effect “yes” or “no” of the different donors on the seed set of the different recipient species.

__Table S5.__ Estimates, standard error, t-value and P-value of the effect of the different 9 donors on each recipient species.

__Table S6.__ Number of seeds produced with 100% foreign pollen treatments for the different recipient species.

__Table S7.__ Phylogenetic signal and significance for all the different traits

__Table S8.__ Procrustes analysis results.

__Figure S1.__ Correlation matrix for all the different traits.

__Figure S2.__ Total amount of pollen found for the different treatments.

__Figure S3.__ Pollen ratios for the different recipient species.

__Figure S4.__ Pollen ratios for the different recipient species by family.

__Figure S5.__ Statistical comparison of pollen ratios by family as pollen donor and recipient.

__Figure S6.__ Violin plot of the reproductive biology of the species.

__Figure S7.__ Grouped effect sizes by family for each recipient species.